Preliminary insights into the use of pottery and culinary pratices at Guijiabao site in southwest China

### 2024-02-16 analysis/paper/paper.Rmd
minor edits

<!--Wild non-ruminant as possible non-ruminant fat origin-->
The possibility of terrestrial animal input for the non-ruminant fat, such as wild boar, cannot be ruled out. The isotopic values of the non-ruminant sample align with those of reference fats from modern wild boars in western Japan [@Lucquin2016] and in the southeast Tibetan Plateau [@Zhang2022]. Although horses could be another non-ruminant fat source, it is less likely to have been part of the local diet in southwest China during the Neolithic period, given the only sporadic occurrences in the south [@Yuan2006]. Domesticated horses probably did not appear in southwest China until the Dian culture period (4th century BC to 1st century AD) in eastern Yunnan [@Lei2009; @Chiou2004], a period approximately 3000 years later than the Neolithic Guijiabao. Ongoing research on faunal remains at Guijiabao will provide more details for further discussion.

@@

# Conclusion
This paper presents a preliminary result investigating pottery use and culinary practices at Guijiabao in the Yanyuan area, middle Hengduan mountain range, through organic residue analyses. Despite the small sample size, our results reveal a relatively diverse range of food categories cooked in pottery during the Neolithic period, including ruminant adipose fats, plants such as broomcorn millet, and non-ruminant sources. In the historical period, mixtures of ruminant adipose fats and non-ruminant fats were identified. Non-ruminant fats could have originated from C3 plants, pigs, or freshwater resources.

>>>>

<!--Wild non-ruminant as possible non-ruminant fat origin-->
The possibility of terrestrial animal input for the non-ruminant fat, such as wild boar, cannot be ruled out. The isotopic values of the non-ruminant sample align with those of reference fats from modern wild boars in western Japan [@Lucquin2016] and in the southeast Tibetan Plateau [@Zhang2022]. Although horses could be another non-ruminant fat source, it is less likely to have been part of the local diet in southwest China during the Neolithic period, given the only sporadic occurrences in the south [@Zhang2022]. Domesticated horses probably did not appear in southwest China until the Dian culture period (4th century BC to 1st century AD) in eastern Yunnan [@Lei2009; @Chiou2004], a period approximately 3000 years later than the Neolithic Guijiabao. Ongoing research on faunal remains at Guijiabao will provide more details for further discussion.

@@

# Conclusion
This paper presents a preliminary result investigating pottery use and culinary practices at Guijiabao in the Yanyuan area, middle Hengduan mountain range, through organic residue analyses. Despite the small sample size, our results reveal a relatively diverse range of food categories cooked in pottery during the Neolithic period, including ruminant adipose fats, plants such as broomcorn millet, and non-ruminant sources. In the historical period, mixtures of ruminant adipose fats and non-ruminant fats were identified. Non-ruminant fats could have originated from C~3~ plants, pigs, or freshwater resources.
